\section{Conclusion}\label{sec:conclude}
Declarative style analysis has become a widely used analysis technique, but it
cannot deal with multilingual programs actively developed in various domains.
In this paper, we propose an extension methodology for a declarative style
analyzer that supports multiple target languages to analyze multilingual
programs.
First, we create a merged database consisting of multiple logical language
spaces. 
Each language space stores facts transformed from source code written in its
corrensponding language.
Then, we define additional language-interoperation rules to derive new facts
from facts across language boundaries.
Our prototype implementation \ours built on top of Code QL tracks dataflows
successfully over language boundaries in both Java-C programs and Python-C
programs.
Using \ours, we found 33 true bugs and vulnerabilities from real-world
applications, 12 of which are from applications that \lees failed to analyze.
We believe that our approach is applicable to various multilingual programs
beyond Java-C and Python-C, and to various types of analyses.

%In this paper, we suggested the methodology of extending a declarative style
%anlyzer that targets multiple languages into a analyzer that can analyze the
%multilingual programs. First we create the database of fact that can be divided
%into multiple logical language spaces, where each language space consists of
%original facts. Then we define additional language-interoperation rules that
%will generate the facts that reflect interoperation semantics between languages.
%Finally, we perform the original query to obtain the analysis result.
%Based on this approach, we build our prototype implementation \ours on top of
%CodeQL.  Using \ours, we could sucessfully perform dataflow analysis on various
%benchmark suites of JNI programs and extension module programs.  Also, we find
%33 true bugs and vulnerabilities from real-world applications, 12 of which are
%from applications that \lees failed to detect.  We believe that our approach is
%applicable to analysis for other kinds of multilingual programs beyond Java-C
%or Python-C, and to various types of analysi beyond dataflow analysis.
