\section{Introduction}

Declarative static analysis is a paradigm of static analysis for detecting bugs
and security vulnerabilities in programs.
In the paradigm, analysts define rules for ``what'' the analysis results in,
rather than algorithms for ``how'' results are calculated in the analysis.
A declarative static analysis is typically performed in three steps.
First, it transforms a program source code to a database containing
\textit{facts} that are tuples of values. 
Second, it derives new facts from known facts in the database by applying {\it
rules}. 
Each rule specifies a set of base facts and a derived fact that will be added
to the database if satisfying all the base facts.
%Each rule specifies that when a certain set of facts satisfying a certain
%condition is included in the database, then a new certain fact also can be
%added to the database.
Finally, the analysis applies {\it queries} for specific properties via a query
system and produces the results of the queries as its analysis results.

Thanks to its declarative characteristic, declarative static analysis has
become one of the widely-used techniques.
Datalog has become a promising domain specific language for static
analysis~\cite{doop, allen2015D, allen2015stagedD, alpuente2010D, dawson1996D,
naik2006D, reps1994D, smaragdakis2014D, whaley2005D, scholz2016}, by
alleviating challenges and burdens in crafting a static program analyzer.
For exameple, \doop~\cite{doop}, a static program analyzer using Datalog, shows
remarkable scalability for the Java points-to analysis, and the scalability
comes from the modularity and declarativeness of Datalog~\cite{doopWorkshop}.
%Researchers have also improved static analysis backends using declarative
%language engines~\cite{whaley2005D, hoder2011muz, souffle, madsen2016}.  
A declarative semantic code analysis engine maintained by Github, called
CodeQL~\cite{codeql}, is scalable enough in the sense that it can detect
security vulnerabilities in a Java programs with over millions of lines.
%Avgustinov et al.~\cite{ql2016} proposed QL, a declarative and object-oriented
%query language that is compiled into Datalog and runs on a relational database. 
%CodeQL~\cite{codeql} is a declarative semantic code analysis engine based on
%QL, maintained by Github. 
%It performs dataflow analysis to detect security vulnerabilities in Java
%programs, and is scalable to large-size programs with millions of lines of
%code. 
Recently, Glean~\cite{glean}, an experimental declarative query system, has
been developed by Meta, on which users can query information about code
structures on a large-scale codebase for JavaScript and Hack programs.
%Recently, Glean~\cite{glean}, an experimental declarative query system, has
%been developed by Meta, on which users can query information about code
%structures for JavaScript and Hack programs.

Declarative static analyzers have broadened their analysis targets from a
single programming language to diverse programming languages. 
To support a new language, one should consider
the three steps of declarative static analysis. Because the query
system is independent of analysis target languages, one can reuse the
existing query system for the new language. Therefore, the first step
is to define a new schema of the database containing facts for the new
language and implement a
front-end component that transforms programs written in the new language into facts.
Then the next step is to define new language-specific rules for the new language.  With these
modifications, declarative analyzers can analyze programs written in the new
language, using the existing queries and the query system.  \doop~currently
supports Python program analysis to detect tensor shape mismatching bugs in
TensorFlow-based Python ML models~\cite{lagouvardos2020static}. 
CodeQL now can track dataflows not only in Java but also in C++, C\#,
JavaScript, Ruby, and Python programs.  Furthermore, Glean plans to support
diverse programming languages, including Python, Java, C++, Rust, and Haskell.

While the analyzers support multiple programming languages, they do not
directly support the analysis of \textit{multilingual programs} written in two
or more programming languages. 
Multilingual programs are now widely developed in various application
domains~\cite{kochhar2016large, mergendahlcross}. However, multilingual
programs are often vulnerable to bugs or security issues more than monolingual
programs. A large-scale study on the code quality of multilingual
programs~\cite{kochhar2016large} reported that using multiple languages
together correlates with higher error-proneness. Moreover, Grichi et
al.~\cite{grichi2020impact} showed that two or three times more bugs and
security issues had been reported in the language interoperation than in the
intraoperation in widely-used open-source JNI programs such as OpenJ9 and VLC.

In this paper, we propose a systematic methodology that extends a
declarative static analyzer supporting multiple languages to support
multilingual program analysis as well. Our goal is to maximize the reuse of
already existing components. First, it generates a merged database of facts that can
be separated into multiple logical language spaces.  Each language space
consists of original facts from its corresponding language database, and existing
language-specific rules derive new facts from the facts in the corresponding
language space. Then, to handle the language interoperation semantics in
multilingual programs, we define language-interoperation rules referring to
the language interoperation semantics. The language-interoperation rules derive new
facts from the facts across language spaces. The extensions enable the query system
to extract facts in multilingual programs, taking the language
interoperation semantics into account.
Finally, the same query is evaluated under the same query system to get
analysis results.

To evaluate the practicality of our approach, we develop a proof-of-concept
declarative static analyzer, called \ours, for multilingual programs by
extending CodeQL. \ours tracks dataflows across language boundaries for two
types of
multilingual programs: Java-C programs written in Java and C and Python-C
programs written in Python and C. The extension is simple enough in that
it requires only a few lines of automated modifications of CodeQL and additional
language-interoperation rules. We also implement a bug checker on top of
\ours to detect dataflow-related interoperation bugs in real-world Java-C programs.
The evaluation shows that our tool successfully tracks dataflows across
Java-C and Python-C language boundaries. 
The evaluation shows that \ours is scalable enough in the sense that it takes
only about 12 minutes to analyze multilingual programs with three million
lines, which the state-of-the-art analyzer fails to analyze in eight hours.
Also, it detects 33 genuine interoperation bugs in real-world Java-C programs,
including 12 new bugs that the existing analyzer could not detect due to the
lack of its scalability.

The contributions of this paper are as follows:
\begin{itemize}
\item We propose a systematic methodology that extends a declarative static analyzer
supporting multiple languages to support multilingual program analysis.

\item We implement a proof-of-concept declarative static dataflow analyzer,
called \ours, for two types of multilingual programs, Java-C and Python-C, by
extending CodeQL.

\item We show that \ours can sucessfully detect dataflow-related bugs at
language boundaries of real-world multilingual programs, including new bugs
that the state-of-the-art analyzers could not detect due to the lack of the
scalability.

\end{itemize}
